# Introduction to Forecasting

## What is Forecasting?

Forecasting is the process of making predictions about future events based on past and present data. In the context of this course, we'll explore various statistical and machine learning methods for forecasting time series data.

## Types of Forecasting

There are several approaches to forecasting:

1. **Qualitative Methods**: Based on expert judgment and opinions
2. **Quantitative Methods**: Based on mathematical models and historical data
3. **Hybrid Methods**: Combining qualitative and quantitative approaches

## Applications

Forecasting has wide applications across various domains:

- **Economics**: GDP growth, inflation rates, unemployment
- **Finance**: Stock prices, exchange rates, commodity prices
- **Business**: Sales, demand, revenue
- **Weather**: Temperature, precipitation, severe weather events
- **Energy**: Electricity demand, renewable energy production

## Course Objectives

By the end of this course, you should be able to:

- Understand the fundamentals of time series data
- Apply various forecasting methods
- Evaluate and compare forecasting models
- Implement forecasting solutions using R

## Getting Started

To begin your forecasting journey, you'll need:

```{r eval=FALSE}
# Essential R packages for forecasting
install.packages(c("forecast", "fpp3", "tsibble", "fable", "ggplot2"))
```

Let's start with a simple example:

```{r simple-forecast, eval=FALSE}
# Load required libraries
library(forecast)
library(ggplot2)

# Create a simple time series
data <- c(10, 12, 13, 12, 14, 16, 18, 17, 19, 20, 22, 21)
ts_data <- ts(data, frequency = 12, start = c(2023, 1))

# Simple forecast
forecast_result <- naive(ts_data, h = 6)

# Plot the forecast
autoplot(forecast_result) +
  ggtitle("Simple Naive Forecast") +
  xlab("Time") +
  ylab("Value")
```

This chapter provides the foundation for understanding forecasting concepts that we'll build upon in subsequent chapters.