# Time Series Fundamentals

## Understanding Time Series Data

A time series is a sequence of observations taken sequentially in time. Understanding the characteristics of time series data is crucial for effective forecasting.

## Components of Time Series

Time series typically consist of several components:

### Trend
The long-term movement or direction in the data.

### Seasonality
Regular patterns that repeat over known, fixed periods (e.g., monthly, quarterly, yearly).

### Cyclical Patterns
Fluctuations that occur over irregular periods, often related to business cycles.

### Irregular/Random Component
Random fluctuations that cannot be explained by trend, seasonal, or cyclical components.

## Time Series Decomposition

Decomposition helps us understand the underlying structure of our data:

```{r decomposition-example, eval=FALSE}
# Example of time series decomposition
library(forecast)

# Decompose the time series
decomposed <- decompose(ts_data)

# Plot the decomposition
plot(decomposed)
```

## Stationarity

A stationary time series has:
- Constant mean
- Constant variance
- Autocovariance depends only on lag

### Testing for Stationarity

```{r stationarity-test, eval=FALSE}
# Augmented Dickey-Fuller test
library(tseries)
adf.test(ts_data)

# KPSS test
kpss.test(ts_data)
```

## Data Preprocessing

Before forecasting, we often need to:

1. **Handle missing values**
2. **Remove outliers**
3. **Transform data** (log, Box-Cox, differencing)
4. **Make series stationary**

### Differencing

```{r differencing, eval=FALSE}
# First difference
diff_data <- diff(ts_data)

# Seasonal differencing
seasonal_diff <- diff(ts_data, lag = 12)
```

## Visualization

Effective visualization is key to understanding time series patterns:

```{r visualization, eval=FALSE}
library(ggplot2)
library(dplyr)

# Basic time series plot
autoplot(ts_data) +
  ggtitle("Time Series Plot") +
  xlab("Time") +
  ylab("Value")

# ACF and PACF plots
acf(ts_data)
pacf(ts_data)
```

Understanding these fundamentals will prepare you for applying specific forecasting methods in the next chapter.