# Forecasting Methods

## Overview

This chapter covers the main forecasting methods used in practice, from simple approaches to advanced techniques.

## Simple Methods

### Mean Method
Forecasts equal the mean of historical observations.

### Naive Method
Forecasts equal the last observed value.

### Seasonal Naive
Forecasts equal the last observed value from the same season.

### Drift Method
Forecasts equal last value plus average change.

```{r simple-methods, eval=FALSE}
library(forecast)

# Apply simple methods
mean_forecast <- meanf(ts_data, h = 12)
naive_forecast <- naive(ts_data, h = 12)
seasonal_naive <- snaive(ts_data, h = 12)
drift_forecast <- rwf(ts_data, h = 12, drift = TRUE)
```

## Exponential Smoothing

Exponential smoothing methods assign exponentially decreasing weights to past observations.

### Simple Exponential Smoothing
For data with no trend or seasonal pattern.

### Holt's Method
For data with trend but no seasonality.

### Holt-Winters Method
For data with both trend and seasonality.

```{r exponential-smoothing, eval=FALSE}
# Simple exponential smoothing
ses_forecast <- ses(ts_data, h = 12)

# Holt's method
holt_forecast <- holt(ts_data, h = 12)

# Holt-Winters
hw_forecast <- hw(ts_data, h = 12)
```

## ARIMA Models

ARIMA (AutoRegressive Integrated Moving Average) models are powerful tools for forecasting.

### Components
- **AR (p)**: Autoregressive terms
- **I (d)**: Degree of differencing
- **MA (q)**: Moving average terms

### Model Selection

```{r arima-models, eval=FALSE}
# Automatic ARIMA model selection
auto_arima <- auto.arima(ts_data)

# Manual ARIMA specification
manual_arima <- Arima(ts_data, order = c(1, 1, 1))

# Forecast
arima_forecast <- forecast(auto_arima, h = 12)
```

## State Space Models

State space models provide a unified framework for time series analysis.

### ETS Models
Error, Trend, and Seasonal components.

```{r ets-models, eval=FALSE}
# ETS model
ets_model <- ets(ts_data)
ets_forecast <- forecast(ets_model, h = 12)
```

## Machine Learning Approaches

Modern forecasting increasingly uses machine learning methods.

### Neural Networks

```{r neural-networks, eval=FALSE}
# Neural network forecast
library(forecast)
nn_forecast <- nnetar(ts_data)
nn_prediction <- forecast(nn_forecast, h = 12)
```

### Other ML Methods
- Random Forest
- Support Vector Machines
- Gradient Boosting

## Model Combination

Combining forecasts from multiple models often improves accuracy.

```{r model-combination, eval=FALSE}
# Simple average of forecasts
combined_forecast <- (arima_forecast$mean + ets_forecast$mean) / 2
```

## Practical Considerations

When choosing a forecasting method, consider:

1. **Data characteristics** (trend, seasonality, noise)
2. **Forecast horizon** (short-term vs. long-term)
3. **Available data** (length and quality)
4. **Computational resources**
5. **Interpretability requirements**

The next chapter will cover how to evaluate and compare these different methods.
