---
title: "The RockCast: University Forecasting Project"
author: "Your Name"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib]
biblio-style: apalike
link-citations: yes
description: "This is a bookdown project for the University Forecasting course."
---

# Preface {-}

Welcome to The RockCast, a comprehensive guide to forecasting methods and applications. This book serves as both a learning resource and reference for the University Forecasting Project.

This book was created using [bookdown](https://bookdown.org/), which makes it easy to write books and long-form articles/reports with R Markdown.

## How to use this book {-}

Each chapter in this book covers different aspects of forecasting:

- **Chapter 1**: Introduction to Forecasting
- **Chapter 2**: Time Series Fundamentals
- **Chapter 3**: Forecasting Methods
- **Chapter 4**: Model Evaluation

You can navigate through the book using the table of contents on the left, or use the search function to find specific topics.

## Prerequisites {-}

This book assumes basic knowledge of:

- Statistics and probability
- R programming
- Time series concepts

## Software Information {-}

This book was built with:

```{r eval=FALSE}
# R packages used
library(bookdown)
library(rmarkdown)
library(knitr)
```

```{r include=FALSE}
# Automatically create a bibliography database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```